---
title: 'HW1: Basic Data Objects'
author: "Firat Seydim"
subtitle: Stat 33A, Fall 2021
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

-----

## 1) Vectors [8 pts]

Create vectors and factors for the columns in the data table displayed above, according to the following data types. If there are missing values, codify them as `NA`. 

- `number`: integer vector

```{r}
number <- c(30L,35L,23L,9L,11L,27L,34L,6L,3L,0L)
typeof(number)
```

- `player`: character vector

```{r}
player <- c("Stephen Curry","Kevin Durant", "Draymond Green","Andre Iguadala", "Klay Thompson","Zaza Pachulia", "Shaun Livingston","Nick Young", "David West","Patrick McCaw")
typeof(player)
```

- `position`: factor

```{r}
position <- factor(c( levels = "PG","PF","PF","SF","SG","C","PG","SG","C","SG"))
typeof(position)
```

- `height`: character vector

```{r}
height <- c("6-3","6-9","6-7","6-6","6-7","6-11","6-7","6-7","6-9","6-7")
typeof(height)
```

- `weight`: double (i.e. real) vector

```{r}
weight <- c(190,240,230,215,215,270,192,210,250,185)
typeof(weight)
```

- `birthdate`: character vector

```{r}
birthdate <- c("March 14, 1988","September 29 1988", "March 4, 1990", "January 29, 1984", "February 8, 1990", "February 10, 1984", "September 11, 1985", "June 1, 1985", "August 29, 1980", "October 25, 1995")
typeof(birthdate)
```

- `experience`: integer vector

```{r}
experience <- c(8L,10L,5L,13L,6L,14L,12L,10L,14L,1L)
typeof(experience)
```

- `college`: character vector

```{r}
college <- c("Davidson College","University of Texas", "Michigan State University", "University of Arizona","Washington State University",NA,NA,"University of Southern California", "Xavier University","University of Nevada")
typeof(college)
```



------


## 2) Vector Subsetting (i.e. indexing) [10 pts]

Write R commands (just one single-line commands)---displaying the output---that
answer the following questions:

a) What is the weight of the tallest player?

```{r}
weight[which.max(height == "6-11")]
```

b) What is the college of the player that has a height of `6-6`?

```{r}
college[which(height == "6-6")]
```

c) What is the position of the player with more years of experience?
_Hint_: the `which.max()` function is your friend.

```{r}
position[which.max(experience)]
```

d) What is the number of the lightest player? _Hint_: the `which.min()` 
function is your friend.

```{r}
number[which.min(weight)]
```

e) What is the average height for those players with more than 5 years of
experience?

```{r}
##SKIP
```

f) How many players have a weight larger than the average (i.e. mean) weight?

```{r}
length(which(weight > mean(weight)))
```

g) How many players have between 9 and 12 years of experience (inclusive)?

```{r}
length(which(experience >= 9 & experience <= 12))
```

h) What is the mean years of experience of _Shooting Guard_ (`SG`) players?

```{r} 
mean(experience[position == "SG"])
```

i) What is the median weight of those players with a position different of _Center_ (`C`)?

```{r}
median(weight[which(position != "C")])
```

j) What is the first quartile (i.e. bottom 25th percentile) of years of
experience among _Power Forwards_ (`PF`) and _Shooting Guards_ (`SG`).
_Hint_: the `quantile()` function is your friend.

```{r}
quantile(experience[which( position == "PF" | position == "SG")], prob=c(.25,.5,.75))

```


-----


## 3) List for GSW [5 pts]

Use the vectors created in the previous section to create the following list `gsw`:

```{r}
gsw <-list(player = player,
           number = number,
           position = position,
           weight = weight,
           experience = experience)
```

Use the list `gsw` to write R commands---displaying the output---that answer the following questions (use only the list `gsw`, NOT the individual vectors):

a) What is the number of the heaviest player?


```{r}
gsw$number[which.max(gsw$weight)]

```

b) What is the position of the player with the least amount of experience?

```{r}
gsw$position[which.min(gsw$experience)]
```

c) How many players have less than 8 or more than 11 years of experience?

```{r}
length((which(gsw$experience < 8 | gsw$experience > 11)))
```

d) What is the third quartile (i.e. bottom 75th percentile) of years of 
experience among _Power Forwards_ (`PF`) and _Shooting Guards_ (`SG`).
Tip: the function `quantile()` is your friend.

```{r}
quantile(gsw$experience[which(gsw$position == "PF" && gsw$position == "SG")], probs = c(.25,.5,.75,1))


```

e) What is the name of the player whose weight is furthest from the average 
weight (of all players)? Tip: the function `which.max()` is your friend.

```{r}
gsw$player[which.max(gsw$weight > mean(gsw$weight))]

```


-----


## 4) More lists [6 pts]

Consider the following list:

```{r}
hp <- list(
  first = 'Harry',
  last = 'Potter',
  courses = c('Potions', 'Enchantments', 'Spells'),
  sport = 'quidditch',
  age = 18L,
  gpa = 3.9
)
```


Write R commands---displaying the output---to answer the following questions:

a) What is the class of `hp`?

```{r}
typeof(hp)
```


b) How many elements are in `hp`?

```{r}
length(hp)
```


c) What is the length of `courses`?

```{r}
length(hp$courses)
```


d) What is the data type of the element `age`?

```{r}
typeof(hp$age)
```


e) What is the data type of the element `gpa`?

```{r}
typeof(hp$gpa)
```


f) If you combine `age` and `gpa` in a new vector, what is the data type of 
this vector?

```{r}
new_hp <- c(hp$age,hp$gpa)
new_hp
typeof(new_hp)
```


-----


## 5) Technical Questions [4 pts]

a) Explain why the following command returns `2`?

```r
1 + TRUE
```
In this case R interprets logical value as 1 which returns 2. 


b) Explain why the following command returns `TRUE`?

```r
"1" == 1
```
That's because this command is logical. Indeed, 1 equals to 1. 

c) Explain why the following command returns `FALSE`?

```r
"-2" > 0
```
This command returns the "FALSE" because, numerically it's not correct. -2 is not larger than 0. If it was -2 < 0 then it would have returned  "TRUE"

d) Explain why the following command returns `TRUE`?

```r
(10 <= 5) >= 0
```
As long as the left side of the statement is greater than or equal to 0, the command will return TRUE. However, when the left side of the command equals less than 0 then it will return a FALSE.


-----


## 6) Explanations [10 pts]

Consider the following vector `lord`:

```{r}
lord <- c('v', 'o', 'l', 'd', 'e', 'm', 'o', 'r', 't')
```



Run the following commands and explain what's happening in each of them (in terms of subsetting, coercion, recycling, vectorization, etc):

a) `lord[TRUE]`

```{r}
lord[TRUE]

```
inputting logical term "TRUE" in a vector will give out exactly the vector. It's the dame same inputting print(lord).

b) `lord[length(lord) + 1]`

```{r}
lord[length(lord) + 1]
```
It means that there is a missing value. It might have try to do vectorization but it won't work as it's character.

c) `lord[seq(from = length(lord), to = 1, by = -2)]`

```{r}
lord[seq(from = length(lord), to = 1, by = -2)]
```
We are commanding to start from the end of the vector data and show output skipping two every time. We are basically generating a sequance.


d) `lord[lord == "o"]`

```{r}
lord[lord == "o"]
```
We are commanding it to output character "o"'s from the vector. In which there are 2 zeros. 

e) `lord[lord != "e" & lord != "o"]`

```{r}
lord[lord != "e" & lord != "o"]
```
With the "!=" command we are saying do not include "e" and "o" in our output. That's why it skips e and o. vo

f) `lord[lord %in% c('a', 'e', 'i', 'o', 'u')]`

```{r}
lord[lord %in% c('a', 'e', 'i', 'o', 'u')]
```
We are using a comparison operator to check if the lord vector has any of the characters in the command. Thus, it outputs only o's and e.

g) `toupper(lord[!(lord %in% c('a', 'e', 'i', 'o', 'u'))])`

```{r}
toupper(lord[!(lord %in% c('a', 'e', 'i', 'o', 'u'))])
```
toupper capitalizes every character in the output. This time we don't want any of the characters written on the c vector. That's why we have used "!". Bacically, we are saying that if there are any of these characters don't include it in the output and for it to capitilize the output. 

h) `paste(lord, collapse = '')`

```{r}
paste(lord, collapse = '')
```
As the argument is a vector. Paste converts the arguments to character strings. Also, we used collapse to remove "".

i) `lord[is.na(lord)]`

```{r}
lord[is.na(lord)]
```
We are basing setting vector to NA. Is.Na indicates which elements are missing. In this case there arent any missing elements, thus; character(0)

j) `sum(!is.na(lord))`

```{r}
sum(!is.na(lord))
```
This time we are summing all of the elements that are NOT missing in the vector and counting each element. Since we used ! and is.na in the same argument and commanded it to sum the characters, it will output 9. 

-----


## 7) Matrices [7 pts]

You can take the vector `lord` and pass it to `matrix()` to create various
kinds of matrices. For example:

```{r}
vol <- matrix(lord, nrow = 3, ncol = 3)
vol
```

and

```{r echo = FALSE, comment = ''}
dem <- matrix(lord[4:9], nrow = 3, ncol = 2)
dem
```

Manipulate the matrix `vol` with bracket notation (i.e. subscripting) to write 
R commands---displaying the output---in order to obtain the following matrices.

a) obtain the following output

```{r}
rbind(vol[1:1,])
```


b) obtain the following output

```{r}
rbind(vol[1:2,2:1])
```


c) obtain the following output

```{r}
cbind(vol[3:1,1:3])
```


d) obtain the following output

```{r}
d <- cbind(vol[1:3,1:2])
dd <- cbind(vol[4:6])
cbind(d,dd)


```


e) obtain the following output

```{r}
e <- cbind(vol[3:1,3:1])
e
```


f) obtain the following output

```{r}
a <- cbind(vol[3:2,3:2], vol[3:2,2:3])
b <- cbind(vol[2:2],vol[1,2],"d","o")
c <- cbind(vol[1,2],vol[2,1])
d <- rbind(b,b)
first_p <- rbind(a,d)
a_reverse <- cbind(vol[2:3,3:2], vol[2:3,2:3])

rbind(first_p,a_reverse)



```


g) obtain the following output

```{r}

g <- cbind(vol[3:1,1:3])
gg <- cbind(vol[1:3,1:3])
ee <- cbind(vol[1:3,3:1])
aa <- cbind(g,e)
bb <- cbind(gg,ee)

rbind(aa,bb)


```
